---
title: Week 11 - Communities in Networks
sort: 5
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
theme_set(theme_graph()) 

library(igraphdata)
data("UKfaculty")

UK <- as_tbl_graph(UKfaculty)

UK <- UK |>
  activate(edges) |>
  filter(edge_is_mutual()) |>
  to_undirected() |>
  simplify() |>
  as_tbl_graph()
```

## Dad Joke

The guy who stole my diary died yesterday.

<span class='fragment'>
My thoughts are with his family.
</span>


## Housekeeping

> - Final project due dates were wrong on website. So sorry!!
> - Make an appointment for Self Assessment Reflection review
> - Submit Exam Questions ASAP
> - Internet technologies demo on Thursday or R labs? Vote?
> - Exam April 16
> - Guest lecture (Scott Feld) + Exam review on April 9

## Plan for today

> - Finish discussion from last week
> - Groups in networks review
> - Discussion questions
> - R

<!---
# Troubled Lands Review

## Questions from Brightspace Assignment

> - How did you play differently in different versions?
> - What in real life is similar to each version of the game?
> - How did you decide whether or not to help others?
> - What does this have to do with social networks and social capital?!


## Debrief

> - What did you think about while playing?
> - How did you feel as you played? Did those feeling change across different versions of the game?
> - When do you help others in real life?

## Debrief

> - Were resources and abilities distributed fairly?
> - How should resources be distributed in the real world?
> - What role did communication play? How did your group make decisions?
> - Did you work to make things equal? Why or why not?
> - What kinds of inequality appeared in the game?
--->

## Community Detection Review

> - What is a component? (Hint: there's often one "giant" component)
> - What is a clique?
> - What is the weakness of cliques as a way of identifying groups?
> - What is a k-plex? (Hint: related to cliques)
> - The goal of most community detection algorithms is high modularity. What does that mean?

## Core and Periphery Review

> - Why do groups often have a core and a periphery?
> - Given some of the other theories we've talked about, how would you expect the periphery to differ from the core?
> - What role does the periphery have in protests (Barberá et al. reading)?
> - What is one way to identify who is in the core or periphery?

## Coreness Example

<div class='col fragment'>
```{r,message=F}
UK |>
  activate(nodes) |>
  mutate(coreness = node_coreness()) |>
ggraph(layout='stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=coreness)) +
  scale_color_viridis_c(name='Coreness')
```

</div>

## Discussion Questions

> - How does betweenness help with community detection?
> - Is there a situation in which those with less knowledge or experience create the core, and the more knowledgeable and experienced create the periphery?
> - Are communities something that naturally exist or are they partly created by methods we use to detect them?
> - Is "slacktivism" actually necessary for large-scale social movements to succeed?
> - How scalable are community detection algorithms?

## Discussion Questions

> - If networks determine access to information and opportunities, is inequality inevitable?
> - Can the same network structure mean different things in different fields, such as social science and biology? If so, is there a common ground between these different kinds of networks?
> - From the perspective of a marketing or advertising professional, would it be better to focus time and resources on the core group (influencers, celebrities, etc.) or the periphery (new people with the potential to share it and start a movement)? Which one is arguably more important? 
> - 

## Visualization Challenge Review

> - Glanced at a handful and they are quite good!
> - Discussion about LLMs and coding
>   - I have been intentionally vague
>   - I want you to learn computational thinking, learn syntax of R
>   - I also want you to learn how to use AI to do cool stuff!
>   - I'm torn about what to expect / encourage
>   - Fundamental knowledge makes AI stuff easier
